---
title: "Matthew Florence Outage EDA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Matthew Florence Outage EDA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The following is a very cursory examination of the Outage data for Hurricanes Matthew and Florence

```{r, include = FALSE}
library(exposur)
library(tigris)
library(tidyverse)
library(sf)
```

```{r, include = FALSE} 
NC_fips <- "37"
NC_counties <- counties(state = NC_fips, cb = TRUE)
NC_tracts <- tracts(state = NC_fips, cb = TRUE)
```

```{r, include = FALSE}
# get dates
matthew_hurdat2  <- get_hurdat2(
  storm_id = "MATTHEW-2016",
  path = "../raw-data/hurdat2"
)
matthew_track <- st_combine(matthew_hurdat2) |>
  st_cast("LINESTRING")
matthew_start <- min(matthew_hurdat2$datetime)
matthew_end <- max(matthew_hurdat2$datetime)

# get outage data
matthew_eaglei <- get_eaglei(
  interval = interval(matthew_start, matthew_end + days(5)),
  path = "../raw-data/eaglei"
)
matthew_outage <- calc_outage(matthew_eaglei, NC_fips)
matthew_outage_sf <- NC_counties |>
  left_join(matthew_outage, by = c("GEOID" = "geoid"))
```

```{r, include = FALSE}
# get dates
florence_hurdat2  <- get_hurdat2(
  storm_id = "FLORENCE-2018",
  path = "../raw-data/hurdat2"
) |>
   filter(datetime > "2018-09-13 12:00:00")
florence_track <- st_combine(florence_hurdat2) |>
  st_cast("LINESTRING")
florence_start <- min(florence_hurdat2$datetime)
florence_end <- max(florence_hurdat2$datetime)

# get outage data
florence_eaglei <- get_eaglei(
  interval = interval(florence_start, florence_end + days(5)),
  path = "../raw-data/eaglei"
)
florence_outage <- calc_outage(florence_eaglei, NC_fips)
florence_outage_sf <- NC_counties |>
  left_join(florence_outage, by = c("GEOID" = "geoid"))
```

## Summary Stats

**Maximum Outage Percentage**: Maximum percentage of customers with no power by county over the course of the storm to 5 days after  
**Average Outage Duration**: Average customer's total duration (days) with no power by county over the course of the storm to 5 days after

```{r, echo = FALSE}

bind_cols(
  Florence = florence_outage$max_pct_out,
  Matthew = matthew_outage$max_pct_out
) |>
  summary() |> 
  knitr::kable(caption = "Maximum Outage Percentage") |>
  kableExtra::kable_styling(full_width = FALSE, position = "float_left")

bind_cols(
  Florence = florence_outage$avg_customer_hrs_out / 24,
  Matthew = matthew_outage$avg_customer_hrs_out / 24
) |>
  summary() |> 
  knitr::kable(caption = "Average Outage Duration") |>
  kableExtra::kable_styling(full_width = FALSE, position = "left")
```

```{r, echo = FALSE}
table <- bind_cols(florence_outage, storm = "Florence") |>
  bind_rows(
    bind_cols(matthew_outage, storm = "Matthew")
  )

outage_pct_boxplot <- ggplot(table) +
  geom_boxplot(aes(y = max_pct_out)) +
  facet_wrap(~ storm) +
  labs(title = "Maximum Outage Percentage")

outage_dur_boxplot <- ggplot(table) +
  geom_boxplot(aes(y = avg_customer_hrs_out / 24 )) +
  facet_wrap(~ storm) +
  labs(title = "Average Outage Duration")

cowplot::plot_grid(outage_pct_boxplot, outage_dur_boxplot)
```

## Correlation
### Matthew
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10}
matthew_prcp <- read_csv("https://raw.githubusercontent.com/ryanzomorrodi/hurricaneexposuredata/master/output/Matthew16/prcp_tr10.csv")
matthew_close <- read_csv("https://raw.githubusercontent.com/ryanzomorrodi/hurricaneexposuredata/master/output/Matthew16/closest_dis_tr10.csv")
matthew_wind <- read_csv("https://raw.githubusercontent.com/ryanzomorrodi/hurricaneexposuredata/master/output/Matthew16/windmodel_tr10.csv")

matthew_prcp <- matthew_prcp |>
  left_join(matthew_close) |>
  mutate(closest_date = as_date(closest_date)) |>
  group_by(geoid) |>
  filter(DATE %within% interval(closest_date - days(1), closest_date + days(2))) |>
  summarize(prcp = sum(PRCPpred))

matthew <- select(matthew_wind, geoid, vmax_sust, vmax_gust) |>
  left_join(matthew_prcp) |>
  left_join(select(matthew_close, geoid, storm_dist)) |>
  mutate(geoid = as.character(geoid)) |>
  mutate(geoid_cnty = str_sub(geoid, 1, 5)) |> 
  left_join(
    x = select(matthew_outage_sf, GEOID, max_pct_out, avg_customer_hrs_out), 
    c("GEOID" = "geoid_cnty")
  )

library(corrr)

matthew_net <- matthew |>
  select(where(is.numeric)) |>
  st_drop_geometry() |>
  correlate(quiet = TRUE, method = "spearman") |>
  mutate(across(where(is.numeric), ~ .x - runif(length(.x), 0, 0.01))) |>
  network_plot() 

matthew_corr <- matthew |>
  select(where(is.numeric)) |>
  st_drop_geometry() |>
  cor(use = "pairwise.complete.obs", method = "spearman") |>
  ggcorrplot::ggcorrplot(type = "upper", colors = c("skyblue1", "white", "indianred2"), lab = TRUE)

cowplot::plot_grid(matthew_net, matthew_corr)
```

### Florence
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10}
florence_prcp <- read_csv("https://raw.githubusercontent.com/ryanzomorrodi/hurricaneexposuredata/master/output/Florence18/prcp_tr10.csv")
florence_close <- read_csv("https://raw.githubusercontent.com/ryanzomorrodi/hurricaneexposuredata/master/output/Florence18/closest_dis_tr10.csv")
florence_wind <- read_csv("https://raw.githubusercontent.com/ryanzomorrodi/hurricaneexposuredata/master/output/Florence18/windmodel_tr10.csv")

florence_prcp <- florence_prcp |>
  left_join(florence_close) |>
  mutate(closest_date = as_date(closest_date)) |>
  group_by(geoid) |>
  filter(DATE %within% interval(closest_date - days(1), closest_date + days(2))) |>
  summarize(prcp = sum(PRCPpred))

florence <- select(florence_wind, geoid, vmax_sust, vmax_gust) |>
  left_join(florence_prcp) |>
  left_join(select(florence_close, geoid, storm_dist)) |>
  mutate(geoid = as.character(geoid)) |>
  mutate(geoid_cnty = str_sub(geoid, 1, 5)) |> 
  left_join(
    x = select(florence_outage_sf, GEOID, max_pct_out, avg_customer_hrs_out), 
    c("GEOID" = "geoid_cnty")
  )

florence_net <- florence |>
  select(where(is.numeric)) |>
  st_drop_geometry() |>
  correlate(quiet = TRUE, method = "spearman") |>
  mutate(across(where(is.numeric), ~ .x - runif(length(.x), 0, 0.01))) |>
  network_plot() 

florence_corr <- florence |>
  select(where(is.numeric)) |>
  st_drop_geometry() |>
  cor(use = "pairwise.complete.obs", method = "spearman") |>
  ggcorrplot::ggcorrplot(type = "upper", colors = c("skyblue1", "white", "indianred2"), lab = TRUE)

cowplot::plot_grid(florence_net, florence_corr)
```

## Maps

### Florence 
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10}
ggplot() +
  geom_sf(data = florence_outage_sf, aes(fill = max_pct_out)) +
  scale_fill_distiller(palette = "Blues", direction = 1) +
  geom_sf(data = florence_track) +
  geom_sf(data = florence_hurdat2, alpha = 0.5, aes(size = wind, color = wind)) +
  scale_size_continuous(range = c(0, 40), limits = c(0, 100)) +
  scale_color_distiller(palette = "BuPu", direction = 1, limits = c(0, 100)) +
  coord_sf(xlim = c(-84, - 75), ylim = c(33, 36.7)) +
  labs(
    title = "Florence Maximum Outage Percentage",
    fill = "Maximum Outage (%)",
    color = "Wind (knots)"
  ) +
  guides(size = "none") +
  theme_void() +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.key.size = unit(0.75, 'cm'),
    legend.position = "bottom",
    legend.title.position = "top",
    legend.key.spacing = unit(0.5, "cm"),
    legend.spacing.x = unit(2, "cm")
  )
```

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10}
ggplot() +
  geom_sf(data = florence_outage_sf, aes(fill = avg_customer_hrs_out / 24)) +
  scale_fill_distiller(palette = "Greens", direction = 1) +
  geom_sf(data = florence_track) +
  geom_sf(data = florence_hurdat2, alpha = 0.5, aes(size = wind, color = wind)) +
  scale_size_continuous(range = c(0, 40), limits = c(0, 100)) +
  scale_color_distiller(palette = "GnBu", direction = 1, limits = c(0, 100)) +
  coord_sf(xlim = c(-84, - 75), ylim = c(33, 36.7)) +
  labs(
    title = "Florence Average Customer Outage Duration",
    fill = "Average Outage (days)",
    color = "Wind (knots)"
  ) +
  guides(size = "none") +
  theme_void() +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.key.size = unit(0.75, 'cm'),
    legend.position = "bottom",
    legend.title.position = "top",
    legend.key.spacing = unit(0.5, "cm"),
    legend.spacing.x = unit(2, "cm")
  )
```

### Matthew
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10}
ggplot() +
  geom_sf(data = matthew_outage_sf, aes(fill = max_pct_out)) +
  scale_fill_distiller(palette = "Blues", direction = 1) +
  geom_sf(data = matthew_track) +
  geom_sf(data = matthew_hurdat2, alpha = 0.5, aes(size = wind, color = wind)) +
  scale_size_continuous(range = c(0, 40), limits = c(0, 100)) +
  scale_color_distiller(palette = "BuPu", direction = 1, limits = c(0, 100)) +
  coord_sf(xlim = c(-84, - 75), ylim = c(33, 36.7)) +
  labs(
    title = "Matthew Maximum Outage Percentage",
    fill = "Maximum Outage (%)",
    color = "Wind (knots)"
  ) +
  guides(size = "none") +
  theme_void() +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.key.size = unit(0.75, 'cm'),
    legend.position = "bottom",
    legend.title.position = "top",
    legend.key.spacing = unit(0.5, "cm"),
    legend.spacing.x = unit(2, "cm")
  )
```

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10}
ggplot() +
  geom_sf(data = matthew_outage_sf, aes(fill = avg_customer_hrs_out / 24)) +
  scale_fill_distiller(palette = "Greens", direction = 1) +
  geom_sf(data = matthew_track) +
  geom_sf(data = matthew_hurdat2, alpha = 0.5, aes(size = wind, color = wind)) +
  scale_size_continuous(range = c(0, 40), limits = c(0, 100)) +
  scale_color_distiller(palette = "GnBu", direction = 1, limits = c(0, 100)) +
  coord_sf(xlim = c(-84, - 75), ylim = c(33, 36.7)) +
  labs(
    title = "Matthew Average Customer Outage Duration",
    fill = "Average Outage (days)",
    color = "Wind (knots)"
  ) +
  guides(size = "none") +
  theme_void() +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.key.size = unit(0.75, 'cm'),
    legend.position = "bottom",
    legend.title.position = "top",
    legend.key.spacing = unit(0.5, "cm"),
    legend.spacing.x = unit(2, "cm")
  )
```